Burning sage: Reversing the curse of dimensionality in the visualisation of high-dimensional data

### 2020-09-23 paper/paper.Rmd
done proofing

```{r pdfsense, fig.cap = "Selected frames of the tour run on the pdfsense data using the default tour display (top), and using the sage display with $R=10$ (bottom). Different underlying physical processes are shown by color and we can see orthogonality between the three groups. The sage display preserves the overall structure while revealing details that are hidden near the center in the default display.", fig.align="center", out.width="0.9\\textwidth"}
frames <- c("001", "051", "078", "092")

@@

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage-088.png}
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage_R1-088.png}
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage_gam20-088.png}
\caption{Selected view of the pollen data in the new sage display, with default settings (left), setting $R=1$ (middle) or $\gamma=20$ (right). We can tune either $\gamma$, $R$ or a combination of the two to reveal the word "EUREKA" near the center of the distribution.}
\label{fig:pollen}
\end{figure}

@@

This paper has introduced the sage tour, which reduces the data piling effects 
that occur when taking low dimensional projections of high dimensional data.
This new technique is easily incorporated into exploratory high dimensional data analysis, and applications shown in Section \ref{sec:application} provide examples of the following tasks:

- classification tasks: the sage display decreases the number of overlapping points between classes and provides better visual separation between classes compared to the regular tour, as shown in the sketches example (Section \ref{sec:appl2})
- clustering tasks: the sage display uncovered clusters that were originally obscured by data piling, while still giving the viewer an accurate assessment of the size of a cluster, and their relative orientation, as shown in the single cell RNA-seq example (Section \ref{sec:appl1})
- shape analysis tasks: the sage display helps us understand structures across multiple dimensions, for example orthogonality between multiple groups, as shown in the pdfsense example (Section \ref{sec:appl3})
- search tasks: the sage display allows to find hidden signal that is concealed by the density of points around the center of the projection, as shown in the pollen example (Section \ref{sec:appl4})

@@

However, the sage display is currently 
naive to information in high dimensional space that is not 
captured by linear projections. Like t-SNE, it would be interesting to explore
transformations of projections where points in the high dimensional space are 
locally dense instead of expanding from the center of the distribution.

@@

The sage tour can be tuned to more aggressively expand the center of the display. 
All of the examples shown had some tuning.  The last example 
demonstrated how points away from the projected center get moved to the edge of the hypersphere as $\gamma$ is increased or $R$ is decreased. The non-linear
transformation can introduce distortions which is a well known problem for any
non-linear dimension reduction technique including t-SNE. However, any distortion
introduced by the sage display are interpretable because it is controlled
by a simple function (Eq. \ref{eq:resc}).

>>>>

```{r pdfsense, fig.cap = "Selected frames of the tour of the pdfsense data using the default tour display (top), and using the sage display with $R=10$ (bottom). Different underlying physical processes are shown by color and we can see orthogonality between the three groups. The sage display preserves the overall structure while revealing details that are hidden near the center in the default display.", fig.align="center", out.width="0.9\\textwidth"}
frames <- c("001", "051", "078", "092")

@@

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage-088.png}
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage_R1-088.png}
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage_gam20-088.png}
\caption{Selected views of the pollen data in the new sage display, with default settings (left), setting $R=1$ (middle) and $\gamma=20$ (right). We can tune either $\gamma$, $R$ or a combination of the two to reveal the word "EUREKA" near the center of the distribution.}
\label{fig:pollen}
\end{figure}

@@

This paper has introduced the sage tour, which reduces the data crowding effects 
that occur when taking low dimensional projections of high dimensional data.
This new technique is easily incorporated into exploratory high dimensional data analysis, and applications shown in Section \ref{sec:application} provide examples of the following tasks:

- clustering: the sage display uncovered clusters that were originally obscured by data piling, while still giving the viewer an accurate assessment of the size of a cluster, and their relative orientation, as shown in the single cell RNA-seq example (Section \ref{sec:appl1})
- classification: the sage display decreases the number of overlapping points between classes and provides better visual separation between classes compared to the regular tour, as shown in the sketches example (Section \ref{sec:appl2})
- shape analysis: the sage display helps us understand structures across multiple dimensions, for example orthogonality between multiple groups, as shown in the pdfsense example (Section \ref{sec:appl3})
- needle discovery: the sage display allows to find hidden signal that is concealed by the density of points around the center of the projection, as shown in the pollen example (Section \ref{sec:appl4})

@@

@@

The tuning parameters can be used to more aggressively expand the center of the display. All of the examples shown had some tuning.  The last example demonstrated how points away from the projected center get moved to the edge of the hypersphere as $\gamma$ is increased or $R$ is decreased. With more center magnification, the  non-linear transformation can introduce distortions, but this is a well known problem for any
non-linear dimension reduction technique including t-SNE. However, unlike t-SNE, any distortion introduced by the sage display is interpretable because it is controlled by a simple function (Eq. \ref{eq:resc}).

### 2020-09-23 paper/paper.tex
done proofing

\caption{Selected frames of the tour run on the pdfsense data using the default tour display (top), and using the sage display with $R=10$ (bottom). Different underlying physical processes are shown by color and we can see orthogonality between the three groups. The sage display preserves the overall structure while revealing details that are hidden near the center in the default display.}\label{fig:pdfsense}
\end{figure}

@@

This paper has introduced the sage tour, which reduces the data piling
effects that occur when taking low dimensional projections of high
dimensional data. This new technique is easily incorporated into
exploratory high dimensional data analysis, and applications shown in
Section \ref{sec:application} provide examples of the following tasks:

\begin{itemize}
\tightlist
\item
  classification tasks: the sage display decreases the number of
  overlapping points between classes and provides better visual
  separation between classes compared to the regular tour, as shown in
  the sketches example (Section \ref{sec:appl2})
\item
  clustering tasks: the sage display uncovered clusters that were
  originally obscured by data piling, while still giving the viewer an
  accurate assessment of the size of a cluster, and their relative
  orientation, as shown in the single cell RNA-seq example (Section
  \ref{sec:appl1})
\item
  shape analysis tasks: the sage display helps us understand structures
  across multiple dimensions, for example orthogonality between multiple
  groups, as shown in the pdfsense example (Section \ref{sec:appl3})
\item
  search tasks: the sage display allows to find hidden signal that is
  concealed by the density of points around the center of the
  projection, as shown in the pollen example (Section \ref{sec:appl4})
\end{itemize}

@@

The sage tour can be tuned to more aggressively expand the center of the
display. All of the examples shown had some tuning. The last example
demonstrated how points away from the projected center get moved to the
edge of the hypersphere as \(\gamma\) is increased or \(R\) is
decreased. The non-linear transformation can introduce distortions which
is a well known problem for any non-linear dimension reduction technique
including t-SNE. However, any distortion introduced by the sage display
are interpretable because it is controlled by a simple function (Eq.
\ref{eq:resc}).

>>>>

\caption{Selected frames of the tour of the pdfsense data using the default tour display (top), and using the sage display with $R=10$ (bottom). Different underlying physical processes are shown by color and we can see orthogonality between the three groups. The sage display preserves the overall structure while revealing details that are hidden near the center in the default display.}\label{fig:pdfsense}
\end{figure}

@@

This paper has introduced the sage tour, which reduces the data crowding
effects that occur when taking low dimensional projections of high
dimensional data. This new technique is easily incorporated into
exploratory high dimensional data analysis, and applications shown in
Section \ref{sec:application} provide examples of the following tasks:

\begin{itemize}
\tightlist
\item
  clustering: the sage display uncovered clusters that were originally
  obscured by data piling, while still giving the viewer an accurate
  assessment of the size of a cluster, and their relative orientation,
  as shown in the single cell RNA-seq example (Section \ref{sec:appl1})
\item
  classification: the sage display decreases the number of overlapping
  points between classes and provides better visual separation between
  classes compared to the regular tour, as shown in the sketches example
  (Section \ref{sec:appl2})
\item
  shape analysis: the sage display helps us understand structures across
  multiple dimensions, for example orthogonality between multiple
  groups, as shown in the pdfsense example (Section \ref{sec:appl3})
\item
  needle discovery: the sage display allows to find hidden signal that
  is concealed by the density of points around the center of the
  projection, as shown in the pollen example (Section \ref{sec:appl4})
\end{itemize}

@@

The tuning parameters can be used to more aggressively expand the center
of the display. All of the examples shown had some tuning. The last
example demonstrated how points away from the projected center get moved
to the edge of the hypersphere as \(\gamma\) is increased or \(R\) is
decreased. With more center magnification, the non-linear transformation
can introduce distortions, but this is a well known problem for any
non-linear dimension reduction technique including t-SNE. However,
unlike t-SNE, any distortion introduced by the sage display is
interpretable because it is controlled by a simple function (Eq.
\ref{eq:resc}).
